Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-18 special-issue-ci/volunteer-clouds-ieee-ci-si.tex

There is, then, space for systems that do not make any of these
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can
% (Paloma) be embedded in?
be embedded in any web-enabled device, including many over which 
the developers have no control.

@@

Other systems, like JET, proposed by Soares et
al. \cite{soares1998get}, also used Java. JET supports
the execution of parallel applications over the Internet and features
a comprehensive statistics collection subsystem, allowing its use for science,
since it yields % (Paloma) Use synonyms, like "yield" or "permit". Repetitive use of "allow" is... repetitive.
 the collection of algorithm analytics by the
experiment designer, unlike the other systems mentioned
above. Other proposals contemporary to JET, such as {\em SuperWeb}
\cite{alexandrov1997superweb} are roughly a translation of classical
distribution techniques, with overly complicated architectures that include
brokers for resource discovery. It is focused
% (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
, however, in solving the
issues involved in browser-based computing and also in making a system
that is able to work across different types of computers, mentioning
SPARCstations and PCs with Windows-95. Interestingly enough, this
paper incorporates a discussion on the economic model, paving the way
for the for-profit distributed computing systems that arose later on. 
On the other hand, the potential for volunteer computing using
browsers was realized
later on \cite{sarmenta-bayanihan} as well as the potential for
mischief of users with code in their hands
% (Paloma) I think this sentence is extremely informal
\cite{sarmenta-sabotagetolerance}. However, these early efforts by
Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
less widespread.

@@

In the work presented in this paper, JavaScript has been used throughout the full
stack, which has the advantage of using the self-same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
when writing code and leverages 
expertise acquired in the library itself.

@@

On the other hand, measuring the performance of the resulting metacomputer
involves understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. But the raw material of
volunteer computing, with regard to
number of users and computation time spent by every one 
in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. Next we describe the
system itself.

As far as we know, this paper presents one of the few experiments which % (Paloma) Change to "which" in order to not repeating "that" all the time
uses computational resources that are as dissimilar as smartphones and
powerful laptops 
or workstation computers in a research center. The methodology used to
gather resources and the algorithms used will be described next, as
well as the results obtained with the initial setup.

%---------------------------------------------------------------
\section{Description of the framework}
\label{sec:description}

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself. This
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}, resulting that, in fact, JavaScript can be
faster than compiled languages such as Scala and, in any case, provides
a performance that is comparable to other languages usually employed  
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be, indeed, faster, but the fact % (Paloma) Too many "fact" words. There was another one in the previous sentence.
 that a single-user
performance is higher is more than offset. This is because the number
of spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets, is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset. This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and them compile it to
JavaScript. For completness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

@@

There are many different ways to validate a framework that intends to
address these challenges. Some of them are included by design: it is
anonymous, since no accounts are needed to participate in the
experiment, it is safe for the user, since we are using the browser
black box, and it is also heterogeneous. We have designed the
experiments so that they show adaptability to different types of
experiments and clients, scalability with the number of clients and
the problem size, and measured performance by comparing it with
a baseline configuration, with which, as indicated above, is
compatible and can be used in a complementary way. Fault tolerance
will be experimented by dropping the server and checking the continuity
of the experiments.

>>>>

There is, then, space for systems that do not make any of these
assumptions, such as the one presented in this paper, which will be
provisionally named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can
% (Paloma) be embedded in?
be embedded in any web-enabled device, including many over which 
developers have no control.

@@

Other systems, like JET, proposed by Soares et
al. \cite{soares1998get}, also used Java. JET supports
the execution of parallel applications over the Internet and features
a comprehensive statistics collection subsystem, enabling its use for science, %(Mario) Was easier to change "allowing"
since it allows % (Paloma) Use synonyms, like "yield" or "permit" 
the collection of algorithm analytics by the
experiment designer, unlike the other systems mentioned
above. Other proposals contemporary to JET, such as {\em SuperWeb}
\cite{alexandrov1997superweb} are roughly a translation of classical
distribution techniques, with overly complicated architectures that include
brokers for resource discovery. SuperWeb is focused
% (Paloma) Careful with starting using plural ("Other proposals", "are") and continuig with singular. If you are referring to "SuperWeb", change "It" for "SuperWeb"
, however, in solving the
issues involved in browser-based computing and also in making a system
that is able to work across different types of computers, mentioning
SPARCstations and PCs with Windows-95. Interestingly enough, this
paper incorporates a discussion on the economic model, paving the way
for the for-profit distributed computing systems that arose later on. 
On the other hand, the potential for volunteer computing using
browsers was realized
later on \cite{sarmenta-bayanihan} as well as the potential of
user misconduct when having access to the source code.
% (Paloma) I think this sentence is extremely informal
\cite{sarmenta-sabotagetolerance}. However, these early efforts by
Sarmenta once again used Java and not JavaScript, making them % (Paloma) making them
less widespread.

@@

In the work presented in this paper, JavaScript has been used throughout the full
stack, which has the advantage of using the self-same lines of code for
writing an evolutionary algorithm for the
desktop or command line and browser clients, both based in the EC
library {\sf NodEO} \cite{DBLP:conf/gecco/GuervosVGES14}. This saves time 
when writing code and leverages 
the expertise acquired
% (Paloma) Not sure if "acquired" should be written before "expertise"... sounds better to me, anyways
 in the library itself.

@@

On the other hand, measuring the performance of the resulting metacomputer
involves understanding the dynamics of this kind of systems. Initial
work was done for peer to peer systems by Stutzbach et
al. \cite{stutzbach2006understanding} and extended to volunteer
computing by Laredo et al. \cite{churn08,laredo2008rcp}. 
%But the raw material of           
%volunteer computing, % (Paloma) Whith regard to?  Maybe it was colloquial/analogy
But some of the essential metrics in volunteer computing like the
number of users or the time spent by every one in the
computation in browser-based volunteer computing experiments, have
only been studied in a limited way in 
\cite{DBLP:journals/gpem/LaredoBGVAGF14} on the basis of a single
run. Studies using volunteer computing platforms such as SETI@home
\cite{javadi2009mining} found out that the Weibull, log-normal and
Gamma distribution 
modeled quite well the availability of resources in several clusters
of that framework; the shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles. This is in concordance with the
results obtained in \cite{agajaj}. Although SETI@home works
at a larger scale in both quantity of volunteers and the scope of projects, 
our initial tests indicate that
this kind of distribution might be the most appropriate to describe
user participation in several different metrics. However, in this
paper we will perform different experiments in different setups and
will try to come up with a more precise model. Next we describe the
system itself.

As far as we know, this paper presents one of the few experiments using % (Paloma) Change to "which" in order to not repeating "that" all the time
computational resources that are as dissimilar as smartphones and
powerful laptops 
or workstation computers in a research center. 
%The methodology used
% (Paloma) I would say "the used methodology"
The algorithms used, as well as the methodology 
for gathering resources will be described next, 
together with the results obtained in this initial setup.
%---------------------------------------------------------------
\section{Description of the framework}
\label{sec:description}

In general, a distributed volunteer-based evolutionary computation
system based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself; which % (Paloma) I recommend the use of "which" here
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact, faster, but the fact that single-user performance is higher (than what?) is over the offset. This is because the number of spontaneous volunteers we can gather at whim, including people using their mobile phones or tablets, is several orders of magnitude higher, which is the objective in this kind of systems." (?)
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and them compile it to
JavaScript. For completness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

@@

There are many different ways to validate a framework that intends to
address these challenges. Some of them are included by design: it is
anonymous, since no accounts are needed to participate in the
experiment, it is safe for the user, since we are using a the browser
black box, and it is also heterogeneous. We have designed the
experiments so that they show adaptability to different types of
experiments and clients, scalability with the number of clients and
the problem size, and also measured its performance comparing it with
a baseline configuration, with which, as indicated above, is
compatible and can be used in a complementary way. Fault tolerance
will be experimented by dropping the server and checking the continuity
of the experiments.
